---
title: "Project 1 - Housing Data Post COVID?"
author: "Mia Alecia Goins, Mak Trnka"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# Then load the package in your R session.
library(ezids)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(results="markup", warning = F, message = F)
options(scientific=T, digits = 3) 
```

# A Peak Under the Hood!
**What does our raw unalterd data look like?**
```{r}
housing = read.csv("C:/Users/flami/OneDrive/Documents/DATS6101/Project 1/RDC_Inventory_Core_Metrics_Metro_History(9-18-26).csv", header =TRUE)

head(housing, n=5)
```
Yikes! let's see if we can clean this up a bit and add our time constraint.\


# Let's Get Relevant!
**Here we will clean up the data, then only select the data from June 2023 to August 2026, the month after the CDC's official declaration of the end of COVID.**

```{r}
#add state column
metro_states = strsplit(housing$cbsa_title, ", ")
metro_states = unlist(metro_states)
states = metro_states[seq(2,length(metro_states),2)]
state = substring(states,1,2)
housing$state_ss = state # it counts DC as a state so there are 51

#cbsa_title is a factor
housing[,3] = as.factor(housing[,3])

# add a month and a year column
housing$month_mm = housing[,1]%%100
housing$year_yyyy = housing[,1]%/%100

# date reformat and time constraint
housing[,1] = as.Date(paste(housing[,1],"01"),'%Y%m%d')
postCOVID_housing = housing[housing$month_date_yyyymm > "2023-05-01",]

#house rank, state, month, and year are factors
postCOVID_housing[,4] = as.factor(postCOVID_housing[,4]) # rank
postCOVID_housing[,48] = as.factor(postCOVID_housing[,48]) # state
postCOVID_housing[,49] = as.factor(postCOVID_housing[,49]) # month
postCOVID_housing[,50] = as.factor(postCOVID_housing[,50]) # year

#make sure its quality
quality_housing = subset(postCOVID_housing, quality_flag==0)
str(quality_housing)

```

Our data has gone from `r nrow(housing)` observations to `r nrow(quality_housing)` after the quality flagging and time constraint being post COVID (June 2023 - August 2026).\

# Basic Stats
**Exploring a little bit about our data**
```{r}
str(quality_housing)
xkablesummary(quality_housing)
```

# What's SMART About the Data?
**Let's explore our SMART Questions!** \

## Question 3
**After COVID (from June 2023 to August 2026) is the average number of price reductions/increases the same across different metro areas, across states?** \

### EDA
It's important to see the variation in our data before constructing our hypotheses. So let's plot the variation in price increase and reduce count. \

```{r}
#maybe add hist
qqnorm(quality_housing$price_increased_count, main="Q-Q plot of Price Increase Count post COVID")
qqline(quality_housing$price_increased_count)

qqnorm(quality_housing$price_reduced_count, main="Q-Q plot of Price Reduced Count post COVID")
qqline(quality_housing$price_reduced_count)
```



Both reductions and increases are right skewed so lets normalize our data then plot again.\

```{r}
library(ggplot2)
#log transform our data
log_housing = quality_housing
log_housing$log_inc_count = log10(log_housing$price_increased_count + 1)
log_housing$log_red_count = log10(log_housing$price_reduced_count + 1)
log_housing$log_list_count = log10(log_housing$total_listing_count + 1)

ggplot(log_housing, aes(x = month_mm, y = log_inc_count)) + 
  geom_boxplot() + 
  geom_boxplot( colour="black", fill="blue", outlier.colour="red", outlier.shape=2, outlier.size=4) +
  labs(title="Boxplot of year vs price increase count post COVID", x="year", y = "log10(price increase count + 1)") 

ggplot(log_housing, aes(x = month_mm, y = log_red_count)) + 
  geom_boxplot() + 
  geom_boxplot( colour="black", fill="red", outlier.colour="red", outlier.shape=2, outlier.size=4) +
  labs(title="Boxplot of year vs price reduced count post COVID", x="year", y = "log10(price reduced count + 1)") 

ggplot(data=log_housing, aes(log_inc_count)) + 
  geom_histogram(breaks=seq(0, 3.5, by = 0.15), 
                 col="green", 
                 fill="blue", 
                 alpha = .7) + # opacity
  labs(x="log10(increase_count)", y="Frequency") +
  labs(title="Histogram of log price increase counts post COVID")

ggplot(data=log_housing, aes(log_red_count)) + 
  geom_histogram(breaks=seq(0, 4.5, by = 0.15), 
                 col="yellow", 
                 fill="red", 
                 alpha = .7) + # opacity
  labs(x="log10(reduced_count)", y="Frequency") +
  labs(title="Histogram of log price reduced counts post COVID")

qqnorm(log_housing$log_inc_count, main="Q-Q plot of log Price Increase Count post COVID")
qqline(log_housing$log_inc_count)

qqnorm(log_housing$log_red_count, main="Q-Q plot of log Price Reduced Count post COVID")
qqline(log_housing$log_red_count)
```
\

Our reduction histogram looks normal but the increase count even after normalizing is still skwed so we cannot conduct an ANOVA test on price increases. So our modified question is:\

**After COVID (from June 2023 to August 2026) is the average number of price reductions the same across different metro areas, across states?**

### Null Hypothesis Construction Household Rank
**Null Hypothesis** $H_0: \mu_1=mu_2=...=\mu_{924}=\mu_{925}$ average log price reduction count is equal across all housing ranks\
**Alternative Hypothesis** $H_1: \mu_1\neq\mu_2\neq...\neq\mu_{924}\neq\mu_{925}$ average log price reduction count is not equal across all housing ranks\
**Alpha** $\alpha=0.05$

### Hypothesis Testing Household Rank
```{r}
anovaLogRank = aov(log_red_count ~ HouseholdRank, data = log_housing)
red_rank_pval = summary(anovaLogRank)[[1]][1,5]
xkabledply(anovaLogRank, title = "Log Price Reduction and Household Rank ANOVA result summary")
```
### Analysis Household Rank
\
Since our P-value is `r red_rank_pval` well below our $\alpha$ (0.05) we reject our null hypothesis and have data that supports our alternative hypothesis that all average log reduction counts are not equal across all house ranks.\

**Post-hoc Tukey HSD Rank** \ 
```{r tukeyHSD}
tukeyAoVRank = TukeyHSD(anovaLogRank)
tukeyAoVRank
```

### Null Hypothesis Contruction State
**Null Hypothesis** $H_0: \mu_1=mu_2=...=\mu_{50}=\mu_{51}$ average log price reduction count is equal across all states (and DC)\
**Alternative Hypothesis** $H_1: \mu_1\neq\mu_2\neq...\neq\mu_{50}\neq\mu_{51}$ average log price reduction count is not equal across all states (including DC)\
**Alpha** $\alpha=0.05$

### Hypothesis Testing State
```{r}
anovaLogState = aov(log_red_count ~ state_ss, data = log_housing)
red_state_pval = summary(anovaLogState)[[1]][1,5]
xkabledply(anovaLogState, title = "Log Price Reduction and State ANOVA result summary")
```

### Analysis State
\
Since our P-value is `r red_state_pval` well below our $\alpha$ (0.05) we reject our null hypothesis and have data that supports our alternative hypothesis that all average log reduction counts are not equal across all states (including DC).\

**Post-hoc Tukey HSD State** \
```{r}
tukeyAoVState = TukeyHSD(anovaLogState)
tukeyAoVState
```


## Question 4
**After COVID (from June 2023 to August 2026) does the average amount of total listings in a metro area fluctuate during the year? Is there an increase during important times (Ex. start of school/summer)?** \

### EDA 
Let's first look at the spread of number of total listings (quantitative). Which have been log10 normalized already.
```{r}
ggplot(log_housing, aes(x = month_mm, y = log_list_count)) + 
  geom_boxplot() + 
  geom_boxplot( colour="black", fill="purple", outlier.colour="red", outlier.shape=2, outlier.size=4) +
  labs(title="Boxplot of Month vs Total Listing Count Post COVID", x="month", y = "log10(total listing count)") 

ggplot(data=log_housing, aes(log_list_count)) + 
  geom_histogram(breaks=seq(0, 6, by = 0.2), 
                 col="orange", 
                 fill="purple", 
                 alpha = .7) + # opacity
  labs(x="log10(total listings)", y="Frequency") +
  labs(title="Histogram of log Total Listing Counts Post COVID")

qqnorm(log_housing$log_list_count, main="Q-Q plot of log total Listing Count Post COVID")
qqline(log_housing$log_list_count)
```

### Null Hypothesis Construction Total Listings
**Null Hypothesis** $H_0: \mu_1=mu_2=...=\mu_{11}=\mu_{12}$ average log total listing count is equal across all 12 months\
**Alternative Hypothesis** $H_1: \mu_1\neq\mu_2\neq...\neq\mu_{11}\neq\mu_{12}$ average log total listing count is not equal across all 12 months\
**Alpha** $\alpha=0.05$

### Hypothesis Testing Total Listings
```{r}
anovaLogList = aov(log_list_count ~ month_mm, data = log_housing)
log_listing_pval = summary(anovaLogList)[[1]][1,5]
xkabledply(anovaLogList, title = "log Total Listing Count and Month ANOVA result summary")
```

### Analysis Total Listings
Since our p-value is `r log_listing_pval` well below our $\alpha$ (0.05) we reject our null hypothesis and have data that supports our alternative hypothesis that the average log total listing count is not equal across all 12 months.

### Post-hoc Tukey HSD Total Listings
```{r}
tukeyAoVList = TukeyHSD(anovaLogList)
tukeyAoVList
```
