---
title: "Project 1 - Housing Data Post COVID?"
author: "Mia Alecia Goins, Mak Trnka"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
library(ezids)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(results="markup", warning = F, message = F)
options(scientific=T, digits = 3) 
```

# A Peak Under the Hood!
**What does our raw unalterd data look like?**
```{r}
housing = read.csv("C:/Users/flami/OneDrive/Documents/DATS6101/Project 1/RDC_Inventory_Core_Metrics_Metro_History(9-18-26).csv", header =TRUE)

str(housing)
```
Yikes! let's see if we can clean this up a bit and add out time constraint.\

# Let's Get Relevant!
**Here we will clean up the date column, reformat cbsa_title, then only select the data from June 2023 to August 2026.**

1. Removing all dates pre-COVID per Metro (Start from June 2023 - August 2026) 
Number of Rows: Raw Dataset (112,850) -> Post-COVID Dataset (36,075)

```{r}
#cbsa_title is a factor
housing[,3] = as.factor(housing[,3])
#time constraint
housing = transform(housing, month_mm =as.integer(month_date_yyyymm%%100))
housing = transform(housing, year_yyyy = as.integer(month_date_yyyymm%/%100))
postCOVID_housing = subset(housing, ((housing$month_mm > 5 & year_yyyy==2023)  | (housing$year_yyyy > 2023)))
```
Our data has gone from over 100,000 observations to a little over 36,000 after the time constraint of Post COVID (June 2023 - August 2026).\

2. Removing all rows that have been flagged
Number of Rows:  Post-COVID Dataset (36,075) -> Post-Covid Data Set without flagged rows (31,827)

```{r}
postCOVID_housing_without_QF <- postCOVID_housing %>%
  filter(is.na(quality_flag) | quality_flag == 0))
```
Now our data have gone from 36,075 to a significantly lower ~9,830, after excluding not just rows that were flagged

3. (OPTIONAL)Removing all metros that had flagged rows 
Number of Rows:  Post-Covid Data Set without flagged rows (31,827) -> Post-Covid Data Set without Metros that had at least one flagged row (9,828)

```{r}
# Find every metro that has at least one flagged month
flagged_metros <- postCOVID_housing %>%
  filter(quality_flag == 1) %>%
  distinct(cbsa_code)

# Keep only metros that were never flagged
postCOVID_housing_without_all_FlaggedMetros <- postCOVID_housing %>%
  filter(!cbsa_code %in% flagged_metros$cbsa_code)
```
Now our data have gone from 31,827 to a an unprecdented 9,828, after excluding not just rows that were flagged, but the metros that had at least one.



# Basic Stats
**Exploring a little bit about our data**
```{r}
xkablesummary(postCOVID_housing)
```

# What's SMART About the Data?
**Some explorations of our SMART Questions** \

## Question 1
**Does the metro area relate to the number of price increases / reductions made by a realtor? Are there more increases in more densely populated areas?** \

### Null Hypothesis Construction

### Hypothesis Testing

### Analysis
