---
title: "Project 1 - Housing Data Post COVID?"
author: "Mia Alecia Goins, Mak Trnka"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# Then load the package in your R session.
library(ezids)
library(ggplot2)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(results="markup", warning = F, message = F)
options(scientific=T, digits = 3) 
```

# A Peak Under the Hood!
**What does our raw unalterd data look like?**
```{r}
housing = read.csv("C:/Users/flami/OneDrive/Documents/DATS6101/Project 1/RDC_Inventory_Core_Metrics_Metro_History(9-18-26).csv", header =TRUE)

#head(housing, n=5)
```
Yikes! let's see if we can clean this up a bit and add out time constraint.\

# Let's Get Relevant!
**Here we will clean up the date column, reformat cbsa_title, then only select the data from June 2023 to August 2026.**

```{r}
#cbsa_title is a factor
housing[,3] = as.factor(housing[,3])

# date reformat and time constraint
housing[,1] = as.Date(paste(housing[,1],"01"),'%Y%m%d')
postCOVID_housing = housing[housing$month_date_yyyymm > "2023-05-01",]

#convert date to date and house rank is a factors
postCOVID_housing[,4] = as.factor(postCOVID_housing[,4]) # rank

#make sure its quality
quality_housing = subset(postCOVID_housing, quality_flag==0)
str(quality_housing)

```

Our data has gone from over 100,000 observations to a little over 31,000 after the time constraint of Post COVID (June 2023 - August 2026).\

# Basic Stats
**Exploring a little bit about our data**
```{r}
xkablesummary(quality_housing)
```

# What's SMART About the Data?
**Some explorations of our SMART Questions** \

## Question 3
**After COVID (from June 2023 to August 2026) did the metro area relate to the number of price increases/reductions made by a Realtor, and are there more increases the lower the house rank?** \

### Plots and Graphs
It's important to see the variation in our data before constructing our hypotheses. So let's plot the variation in price increase and reduce count
```{r}
#split our data up by year
housing2023 = quality_housing[quality_housing$month_date_yyyymm >= "2023-06-01" & quality_housing$month_date_yyyymm <= "2023-12-01", ]
housing2024 = quality_housing[quality_housing$month_date_yyyymm >= "2024-01-01" & quality_housing$month_date_yyyymm <= "2024-12-01", ]
housing2025 = quality_housing[quality_housing$month_date_yyyymm >= "2025-01-01" & quality_housing$month_date_yyyymm <= "2025-12-01", ]
housing2026 = quality_housing[quality_housing$month_date_yyyymm >= "2026-01-01" & quality_housing$month_date_yyyymm <= "2026-12-01", ]
ggplot(housing2023, aes(x=month_date_yyyymm,y= price_increased_count, color=cbsa_code)) +
  geom_point(size = 3) +
  geom_text(aes(label = price_increased_count), 
            position = position_nudge(x = 7, y = 3),
            size = 3, 
            show.legend = FALSE) +
  
  ggtitle("Month vs. Price Increase Count in 2023") +
  xlab("Time (Months)") +
  ylab("Price Increase Count") +
  theme(legend.position = "right")

```

### Null Hypothesis Construction


### Hypothesis Testing

### Analysis

## Question 4
**Does the total amount of listings in a metro area fluctuate during the year? Is there an increase during important times (Ex. start of school/summer)?** \
```{r}
# split data by season
#summer
#winter
#spring
#fall

```

### Null Hypothesis Construction

### Hypothesis Testing

### Analysis
