## Data Preparation for Questions 1 and 2
Questions 1 and 2 compare each metro's median listing price in June 2023 with its price in August 2026. We keep only those two months, place each metro in a size tier using its June 2023 household rank, and calculate one price change per metro.\

```{r}
library(dplyr)
library(ggplot2)
library(ezids) # for xkabledply()

housing <- postCOVID_housing_without_QF

# Convert dates to YYYYMM numbers.
housing$month <- as.integer(
  gsub("-", "", substr(as.character(housing$month_date_yyyymm), 1, 7))
)

housing <- subset(
  housing,
  month %in% c(202306, 202608) &
  !is.na(cbsa_code)
)

# Apply the existing quality-flag rule.
if ("quality_flag" %in% names(housing)) {
  housing <- subset(housing, quality_flag == 0)
}

# Accept either household-rank column name.
if (!("household_rank" %in% names(housing))) {
  housing$household_rank <- housing$HouseholdRank
}

stopifnot("household_rank" %in% names(housing))

housing$household_rank <- as.numeric(as.character(housing$household_rank))

# Treat missing or invalid ranks as missing.
housing$household_rank[
  !is.finite(housing$household_rank) |
  housing$household_rank < 1
] <- NA_real_

# Each metro should appear only once per month.
stopifnot(anyDuplicated(housing[c("cbsa_code", "month")]) == 0)
```

We split the metros into four tiers with about the same number of metros in each, using June 2023 household rank (rank 1 = most households).\

```{r}
housing_start <- housing %>%
  dplyr::filter(month == 202306) %>%
  dplyr::arrange(household_rank, cbsa_code) %>%
  dplyr::transmute(
    cbsa_code,
    cbsa_title,
    household_rank,
    price_2023 = median_listing_price,
    pending_2023 = pending_ratio,
    metro_size_group = factor(
      dplyr::ntile(household_rank, 4),
      levels = 1:4,
      labels = c(
        "Tier 1 (Largest)",
        "Tier 2 (Upper-Mid)",
        "Tier 3 (Lower-Mid)",
        "Tier 4 (Smallest)"
      )
    )
  )

housing_end <- housing %>%
  dplyr::filter(month == 202608) %>%
  dplyr::transmute(
    cbsa_code,
    price_2026 = median_listing_price,
    pending_2026 = pending_ratio
  )

# One row per metro with both endpoints.
metro_endpoints <- housing_start %>%
  dplyr::inner_join(housing_end, by = "cbsa_code") %>%
  dplyr::filter(
    is.finite(price_2023),
    is.finite(price_2026),
    price_2023 > 0,
    price_2026 > 0
  ) %>%
  dplyr::mutate(
    price_change_pct = 100 * (price_2026 - price_2023) / price_2023,
    log_price_change = log(price_2026 / price_2023)
  )

nrow(metro_endpoints)
```


## Question 1
**After COVID (from June 2023 to August 2026), did the average change in median listing price differ across the four metro-size tiers?** \

### EDA
It's important to see the variation in our data before constructing our hypotheses. So let's first plot the percent change in median listing price across all metros.\

```{r}
q1 <- subset(
  metro_endpoints,
  is.finite(log_price_change) &
  !is.na(metro_size_group)
)

stopifnot(anyDuplicated(q1$cbsa_code) == 0)

table(q1$metro_size_group)

q1_colors <- c("blue", "lightblue", "gold", "orange")

ggplot(data = q1, aes(price_change_pct)) +
  geom_histogram(bins = 30,
                 col = "black",
                 fill = "blue",
                 alpha = .7) + # opacity
  labs(x = "% change in median listing price", y = "Frequency") +
  labs(title = "Histogram of % listing-price change post COVID")

qqnorm(q1$price_change_pct, main = "Q-Q plot of % Listing-Price Change post COVID")
qqline(q1$price_change_pct)
```

A percent change can't fall below -100% but has no upper limit, so percent changes tend to be right skewed. So let's log transform our data, using $\log(\text{price}_{2026} / \text{price}_{2023})$, then plot again. For small changes the log change is close to the percent change (a log change of 0.10 is about 10.5%).\

```{r}
ggplot(data = q1, aes(log_price_change)) +
  geom_histogram(bins = 30,
                 col = "black",
                 fill = "purple",
                 alpha = .7) + # opacity
  labs(x = "log(price 2026 / price 2023)", y = "Frequency") +
  labs(title = "Histogram of log listing-price change post COVID")

ggplot(q1, aes(x = metro_size_group, y = log_price_change, fill = metro_size_group)) +
  geom_boxplot(colour = "black", outlier.colour = "red", outlier.shape = 2, outlier.size = 2) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values = q1_colors) +
  labs(title = "Boxplot of metro-size tier vs log listing-price change post COVID",
       x = "Metro-size tier", y = "log(price 2026 / price 2023)") +
  theme(legend.position = "none")

q1_qq_plot <- ggplot(q1, aes(sample = log_price_change, color = metro_size_group)) +
  stat_qq(size = 1.5, alpha = 0.7) +
  stat_qq_line(color = "black") +
  facet_wrap(~ metro_size_group, ncol = 2, scales = "free_y") +
  scale_color_manual(values = q1_colors) +
  labs(title = "Q-Q plots of log listing-price change by tier post COVID",
       x = "Theoretical normal quantiles",
       y = "Log price-change quantiles") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none")

q1_qq_plot

q1_summary <- q1 %>%
  dplyr::group_by(metro_size_group) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean_log_change = mean(log_price_change),
    median_log_change = median(log_price_change),
    sd_log_change = sd(log_price_change),
    avg_pct_change = 100 * (exp(mean(log_price_change)) - 1)
  )

xkabledply(as.data.frame(q1_summary), title = "Log listing-price change by metro-size tier")
```

In the boxplot, black diamonds show the mean of each tier and the lines inside the boxes show the median. The Q-Q plots show the log price changes are roughly normal in all four tiers. Tiers 2 and 3 have a longer upper tail, meaning a few metros had unusually large price gains, but with this many metros per tier the ANOVA holds up well to mild non-normality. So we can conduct an ANOVA test on log listing-price change.\

### Null Hypothesis Construction Metro-Size Tier
**Null Hypothesis** $H_0: \mu_1=\mu_2=\mu_3=\mu_4$ average log listing-price change is equal across all four metro-size tiers\
**Alternative Hypothesis** $H_1:$ at least one tier's average log listing-price change is different\
**Alpha** $\alpha=0.05$

### Hypothesis Testing Metro-Size Tier
```{r}
q1_anova <- aov(log_price_change ~ metro_size_group, data = q1)
q1_pval <- summary(q1_anova)[[1]][1, 5]
xkabledply(q1_anova, title = "Log Listing-Price Change and Metro-Size Tier ANOVA result summary")

# Mean log change per tier, converted to average % change.
q1_means <- tapply(q1$log_price_change, q1$metro_size_group, mean)
q1_pct <- 100 * (exp(q1_means) - 1)
round(q1_pct, 2)

# Check: Welch's ANOVA does not assume the tiers have equal spreads.
q1_welch_pval <- oneway.test(log_price_change ~ metro_size_group, data = q1, var.equal = FALSE)$p.value
q1_welch_pval
```

### Analysis Metro-Size Tier
\
Our p-value is `r format.pval(q1_pval, digits = 3, eps = 0.001)`, which is `r if (q1_pval < 0.05) "below" else "above"` our $\alpha$ (0.05), so we `r if (q1_pval < 0.05) "reject our null hypothesis and have data that supports our alternative hypothesis that average log listing-price change is not equal across all four metro-size tiers" else "fail to reject our null hypothesis. There is not enough evidence that average log listing-price change differs across the four metro-size tiers"`.\

Converted to percentages, `r names(which.max(q1_pct))` had the largest average price change (`r round(max(q1_pct), 1)`%) and `r names(which.min(q1_pct))` had the smallest (`r round(min(q1_pct), 1)`%). Because the tiers have different spreads, we also ran Welch's ANOVA, which gives a p-value of `r format.pval(q1_welch_pval, digits = 3, eps = 0.001)` and `r if ((q1_welch_pval < 0.05) == (q1_pval < 0.05)) "leads to the same conclusion" else "leads to a different conclusion, so this result should be interpreted with caution"`.\

**Post-hoc Tukey HSD Metro-Size Tier** \
```{r}
q1_tukey <- TukeyHSD(q1_anova)
q1_tukey

q1_tukey_table <- as.data.frame(q1_tukey$metro_size_group)
q1_sig_pairs <- rownames(q1_tukey_table)[q1_tukey_table$`p adj` < 0.05]
```

Each row compares two tiers, and pairs with a `p adj` below 0.05 have significantly different average log price changes. Significantly different pairs: `r if (length(q1_sig_pairs) > 0) paste(q1_sig_pairs, collapse = "; ") else "none"`.`r if (q1_pval >= 0.05) " Because the ANOVA was not significant, these comparisons are shown for completeness only." else ""`\


## Question 2
**After COVID (from June 2023 to August 2026), were increases in pending ratios associated with changes in median listing prices across metro areas?** \

**Pending Ratio:** The number of pending listings divided by the number of active listings. It measures the balance between supply (active listings) and demand (pending listings).\
- **Increase in Pending Ratio:** The market got tighter. Demand grew relative to supply, meaning buyers faced more competition and sellers gained more leverage.\
- **Decrease in Pending Ratio:** The market got looser. Supply grew relative to demand, meaning buyers gained more leverage and sellers faced more competition to sell.\

We answer this two ways. An ANOVA compares the **size** of the average log price change between metros whose pending ratio increased and metros whose pending ratio decreased. A chi-square test compares the **direction**, meaning the share of metros whose listing price increased.\

### EDA
Let's first group each metro by the direction of its pending ratio and its listing price. A metro counts as "Increased" if its August 2026 value is higher than its June 2023 value, and "Decreased" otherwise, so "Decreased" also includes the few metros with no change.\

```{r}
q2 <- subset(
  metro_endpoints,
  is.finite(log_price_change) &
  is.finite(pending_2023) &
  is.finite(pending_2026) &
  pending_2023 >= 0 &
  pending_2026 >= 0
)

stopifnot(nrow(q2) > 0)
stopifnot(anyDuplicated(q2$cbsa_code) == 0)

q2$pending_change <- q2$pending_2026 - q2$pending_2023

directions <- c("Increased", "Decreased")

q2$pending_group <- factor(
  ifelse(q2$pending_change > 0, "Increased", "Decreased"),
  levels = directions
)

q2$price_group <- factor(
  ifelse(q2$price_2026 > q2$price_2023, "Increased", "Decreased"),
  levels = directions
)

q2_colors <- c("Increased" = "#eb6834", "Decreased" = "#2a78d6")

table(q2$pending_group)
table(q2$price_group)
```

Next, let's look at the spread of log listing-price change in each pending-ratio group.\

```{r}
ggplot(q2, aes(x = pending_group, y = log_price_change, fill = pending_group)) +
  geom_boxplot(colour = "black", outlier.colour = "red", outlier.shape = 2, outlier.size = 2) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4, colour = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values = q2_colors) +
  labs(title = "Boxplot of pending-ratio direction vs log listing-price change post COVID",
       x = "Change in pending ratio", y = "log(price 2026 / price 2023)") +
  theme(legend.position = "none")

q2_qq_plot <- ggplot(q2, aes(sample = log_price_change, color = pending_group)) +
  stat_qq(size = 1.5, alpha = 0.7) +
  stat_qq_line(color = "black") +
  facet_wrap(~ pending_group, ncol = 2, scales = "free_y") +
  scale_color_manual(values = q2_colors) +
  labs(title = "Q-Q plots of log listing-price change by pending-ratio direction",
       x = "Theoretical normal quantiles",
       y = "Log price-change quantiles") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none")

q2_qq_plot
```

Finally, let's count how many metros fall into each combination of pending-ratio direction and listing-price direction, and show each group's split as a 100% stacked bar.\

```{r}
q2_table <- table(
  q2$pending_group,
  q2$price_group,
  dnn = c("Pending ratio", "Listing price")
)

addmargins(q2_table)

q2_prop <- 100 * prop.table(q2_table, margin = 1)

q2_bar <- as.data.frame(q2_prop)
names(q2_bar) <- c("pending_group", "price_group", "percent")
q2_bar$label <- ifelse(q2_bar$percent > 0, paste0(round(q2_bar$percent, 1), "%"), "")

q2_axis_labels <- paste0(directions, "\nn = ", rowSums(q2_table))

q2_plot <- ggplot(q2_bar, aes(pending_group, percent, fill = price_group)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +
  scale_fill_manual(values = q2_colors, name = "Median listing-price change") +
  scale_x_discrete(labels = q2_axis_labels) +
  scale_y_continuous(breaks = seq(0, 100, 20), labels = paste0(seq(0, 100, 20), "%")) +
  labs(title = "Listing-price direction by pending-ratio direction post COVID",
       x = "Change in pending ratio",
       y = "Percentage of metros",
       caption = "'Decreased' includes unchanged values.") +
  theme_minimal() +
  theme(legend.position = "bottom")

q2_plot
```

### Null Hypothesis Construction Pending-Ratio Direction (ANOVA)
**Null Hypothesis** $H_0: \mu_{increased}=\mu_{decreased}$ average log listing-price change is equal for metros whose pending ratio increased and metros whose pending ratio decreased\
**Alternative Hypothesis** $H_1: \mu_{increased}\neq\mu_{decreased}$ average log listing-price change is different between the two groups\
**Alpha** $\alpha=0.05$

### Hypothesis Testing Pending-Ratio Direction (ANOVA)
```{r}
q2_anova <- aov(log_price_change ~ pending_group, data = q2)
q2_pval <- summary(q2_anova)[[1]][1, 5]
xkabledply(q2_anova, title = "Log Listing-Price Change and Pending-Ratio Direction ANOVA result summary")

# Mean log change per group, converted to average % change.
q2_means <- tapply(q2$log_price_change, q2$pending_group, mean)
q2_pct <- 100 * (exp(q2_means) - 1)
round(q2_pct, 2)

# Check: Welch's ANOVA does not assume the groups have equal spreads.
q2_welch_pval <- oneway.test(log_price_change ~ pending_group, data = q2, var.equal = FALSE)$p.value
q2_welch_pval
```

### Analysis Pending-Ratio Direction (ANOVA)
\
Our p-value is `r format.pval(q2_pval, digits = 3, eps = 0.001)`, which is `r if (q2_pval < 0.05) "below" else "above"` our $\alpha$ (0.05), so we `r if (q2_pval < 0.05) "reject our null hypothesis and have data that supports our alternative hypothesis that average log listing-price change is different between metros whose pending ratio increased and metros whose pending ratio decreased" else "fail to reject our null hypothesis. There is not enough evidence that average log listing-price change differs between the two pending-ratio groups"`.\

Converted to percentages, metros whose pending ratio increased had an average price change of `r round(q2_pct["Increased"], 1)`%, compared with `r round(q2_pct["Decreased"], 1)`% for metros whose pending ratio decreased. Welch's ANOVA gives a p-value of `r format.pval(q2_welch_pval, digits = 3, eps = 0.001)` and `r if ((q2_welch_pval < 0.05) == (q2_pval < 0.05)) "leads to the same conclusion" else "leads to a different conclusion, so this result should be interpreted with caution"`. With only two groups, no post-hoc Tukey test is needed.\

### Null Hypothesis Construction Pending-Ratio Direction (Chi-square)
**Null Hypothesis** $H_0:$ pending-ratio direction and listing-price direction are independent\
**Alternative Hypothesis** $H_1:$ pending-ratio direction and listing-price direction are associated\
**Alpha** $\alpha=0.05$

### Hypothesis Testing Pending-Ratio Direction (Chi-square)
```{r}
# Expected counts if the two directions were independent.
q2_expected <- outer(rowSums(q2_table), colSums(q2_table)) / sum(q2_table)
dimnames(q2_expected) <- dimnames(q2_table)
round(q2_expected, 2)

# Classroom rule: every expected count must be at least 5.
stopifnot(all(q2_expected >= 5))

q2_chi <- chisq.test(q2_table, correct = FALSE)
q2_chi_pval <- q2_chi$p.value
q2_chi
```

### Analysis Pending-Ratio Direction (Chi-square)
\
All expected counts are at least 5, so the chi-square test is valid. Our p-value is `r format.pval(q2_chi_pval, digits = 3, eps = 0.001)`, which is `r if (q2_chi_pval < 0.05) "below" else "above"` our $\alpha$ (0.05), so we `r if (q2_chi_pval < 0.05) "reject our null hypothesis and have data that supports our alternative hypothesis that pending-ratio direction and listing-price direction are associated" else "fail to reject our null hypothesis. There is not enough evidence that pending-ratio direction and listing-price direction are associated"`.\

`r round(q2_prop["Increased", "Increased"], 1)`% of metros whose pending ratio increased saw their median listing price rise, compared with `r round(q2_prop["Decreased", "Increased"], 1)`% of metros whose pending ratio decreased.\

Both tests assume each metro is an independent observation, and they show association, not cause and effect.\
